Fast Algorithms for Intersection of Non-matching Grids Using \plucker coordinates.

### 2016-09-30 paper_intersections.tex
Add figure. 23 intersections in progres

\section{Element Intersections}
\label{sec:element intersecitons}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3D ambient space. 
In particular we are interested in the intersections of 1D-2D, 1D-3D, 2D-2D, 2D-3D pairs of elements. 
The fundamental idea is to compute intersection of 1D-2D simplices using the 
\plucker coordinates and reduce all other cases to this one.

@@

\subsection{Intersection Triangle-Tetrahedron (2D-2D)}
\noteJB{Just note that this algorithm can also be implemented using 1d-2d, but quite complex to not yet fully tested}

@@

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side polygon, $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
So we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of the intersection polygon consists of two parts: calculation of side-tetrahedron ICs (Algorithm \ref{algo:colect_23_ip_triangle})
provides all ICs on the boundary of $S_2$,
calculation of edge-triangle ICs (Algorithm \ref{algo:collect_23_ip_edges}).
\notePE{The following sentence is somehow weird.. I would say that we store ICs $I$; to sort $I$ during the computation, we use the two tables; at the end, $I$ is ordered correctly.}
Correct order of the ICs, stored in the list $I$, is defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. \notePE{The following sentece is a copy of the second sentence is this subsection.} Every side of the intersection polygon 
lies either on a side of $S_2$, or on a face, or on an edge of $S_3$. Let us denote $M_2$ the set of sides of $S_2$ and $M_3$ the set of volume,
faces, edges, and vertices of $S_3$. Every side of polygon that lies on $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$.

@@

  \DontPrintSemicolon
  \SetKw{continue}{continue}
  \KwIn{input data}
  \KwOut{List of ICs on sorted output data}  
  $F_g(:)=-1,\ F_p(:)=-1$\tcp*{Unset links.} 
  \For{side $s$ of $S_2$}{ 
    $L$ = intersection($s, S_3$)\;
    \lIf{ $\abs{L} = 0$}{\continue}
    \lIf{ $\abs{L} = 1$}{
        append $p$ to $J$
        \continue
    }
    \For{$p$ in $L$}{
        \lIf{$p \ne I[-1]$ }{append $p$ to $I$}
        \nllabel{line:m4_object}
        $p$ lies on $x\in M_3$\; 
        \uIf{$p$ is first in $L$}{
            $F_g[x] = p$,\ $F_p[p] = s$\;
        }
        \Else($p$ is the last in $L$){
            $F_g[s] = p,\ F_p[p] = x$\;
            \lIf{$F_g[x] = -1$}{$F_g[x] = p$} \nllabel{line:back_link43}
        }
    }
  }  
  \If{$I[-1] = I[0]$}{
    $x = F_p[I[-1]]$\;
    \If{$F_g[x]]=I[-1]$}{$F_g[x]=I[0]$}
    $F_g[s_2] = I[0]$\;
    remove $I[-1]$
  }
\end{algorithm}

Algorithm \ref{algo:colect_23_ip_triangle} passes through every side $s$ of the triangle $S_2$ and computes the line-tetrahedron intersection $L$.
In the regular case we process each of the two ICs in $L$. The IC $p$ is added to the list $I$ unless the last point is the same. 
This condition is effective just for the first IC in $L$ and merges ICs at the same vertex of $S_2$. Then we identify an object $x\in M_3$ the point
$p$ lies on (line \ref{line:m4_object}). \notePE{I am not sure, if the term 'object' is clear..}

\notePE{Why calling it vertex $v$ instead of IC $p$ here? What is $y$?}
If the vertex $v$ is between sides $s_1$, $s_2$, 
we effectively set connections $F_g[s_1] = v$, $F_g[y]=v$, $F_p[v] = s_2$. The backward temporary link on the line \ref{line:back_link43} 
is used in Algorithm \ref{algo:collect_23_ip_edges} to fix $F_p[p]$ for the second $p$ lying on an edge or a vertex of $S_3$. The condition at the end deals with the case
in which two sides intersect the same object $x\in M_3$.

For the regular intersection $L$, we pass through its ICs, as described above. \notePE{For special cases}, we firstly compute ICs
on the boundary of $S_2$ using the line-tetrahedron intersection algorithm for every side. We store single point intersections into separate list $J$
and skip filling the connection tables. This happens when the side touches $S_3$ at its edge or vertex. 
These ICs will be rediscovered again in Algorithm \ref{algo:collect_23_ip_edges} with better topological information, however
this is not the case if the touched edge $e$ of $S_3$ is coplanar with $S_2$ and the IC is inside of $e$. 
Therefore we keep a separate list $J$ of ICs to deal with this case.

@@

  \lFor{edge $e$ of $S_3$}{$L[e]$ = intersection($e, S_2$)}  \nllabel{line:12edges}
  \For{unmarked edge $e$ of $S_3$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = \edgefaces($e$)
    } 
    \Else($p$ at the vertex $v$ of $S_3$){       
            $(f_0,\ f_1)$ = \vertexfaces($v$,$L$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
            \nllabel{line:mark_edges}
            mark all edges coincident with $p$
    }    
    \uIf{ $p$ is on boundary of $S_2$ }{
         $p$ lies on edge or at vertex $x \in M_3$\;
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$}
         \lElse{$F_g[f_0]=q$}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\;
         $F_g[f_0]=p$\tcp*{overwrite the backlink}
         $F_p[p]=f_1$\;         
    }   
  }

  \lIf{$\abs{I} < 3$}{ return $J$}
  \lElse{ return $I$ sorted according to connectivity in $F_g$ and $F_p$}
\end{algorithm}

@@

Algorithm \ref{algo:collect_23_ip_edges} uses the line-triangle intersection algorithm for the edges of $S_3$ (line \ref{line:12edges}).
The loop produces ICs in the interior of $S_2$ and possibly those ICs with special position on vertex or edge of $S_3$ already computed in Algorithm \ref{algo:colect_23_ip_triangle}. 
Every edge $e$ of $S_3$ is oriented so that the pair of adjacent faces $f_0$, $f_1$ appears in the same order on the intersection polygon when the 
IC on $e$ has a negative sign (see Figure \ref{fig:plucker_products}). The function \emph{edge faces}, used on line \ref{line:edge_faces} and later on,
uses the sign of the intersection to return the pair of faces ordered correctly. 
Similarly the function \emph{vertex faces} (Algorithm \ref{algo:vertex_faces}, described later)
returns a pair of generalized faces (face or edge) possibly adjacent to the IC $L[e]$ at the vertex $v$ of $S_3$.
On line \ref{line:mark_edges} we mark all edges coincident with the vertex, since if they have any other intersection corner they are degenerate and 
not processed anyway.

@@

\begin{algorithm}
  \caption{2d-3d intersection, finish sort of points}  
  \label{algo:order_23}

  \SetKw{To}{to}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $K$ in correct order}  
  $P=(1,\dots, n-1,0)$\tcp{array of sucessors}
  \lFor{ $f$ is face of $S_3$}{$P[ F[1,f] ] = F[0, f]$}
  $i=0$\;  
  \lFor{$n=0$ \To $\abs{I}-1$}{
    $K[n]=I[i]$
    $i=P[i]$
  }    
  \label{algo:order_23}
\end{algorithm}

>>>>

\section{Element Intersections}
\label{sec:element intersecitons}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3D ambient space. 
In particular we address intersection algorithms for 1D-2D, 1D-3D, 2D-3D pairs of elements. We have implemented the case 2D-2D as well however the treatment 
of the special cases is quite technical and not fully completed yet. The fundamental idea is to compute intersection of 1D-2D simplices using the 
\plucker coordinates and reduce all other cases to this one.

@@

%\subsection{Intersection Triangle-Tetrahedron (2D-2D)}
%\noteJB{Just note that this algorithm can also be implemented using 1d-2d, but quite complex to not yet fully tested}

@@

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Algorithm \ref{algo:colect_23_ip_triangle}), 
calculation of edge-triangle ICs (Algorithm \ref{algo:collect_23_ip_edges}), oredering of ICs (Algorithm \ref{algo:order_23}).
The intersection corners appended to the list $I$ as they are computed however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are numbered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interiror with normal pointing to us.

@@

  \DontPrintSemicolon
  \SetKw{continue}{continue}
  \KwIn{input data}
  \KwOut{List of ICs on sorted output data}  
  $F_g(:)=-1,\ F_p(:)=-1$\tcp*{Unset links.} 
  \For{side $s$ of $S_2$}{ \nllabel{line:one side loop}
    $L$ = intersection($s, S_3$)\;
    \lIf{ $\abs{L} = 0$}{\continue}
    \lIf{ $\abs{L} = 1$}{
        append $p$ to $J$
        \continue
    }
    \For{$p$ in $L$}{
%        \lIf{$p \ne I[-1]$ }{append $p$ to $I$}
        \nllabel{line:m4_object}
        $p$ lies on $x\in M_3$\;
        \lIf{$p$ lies at vertex $v$ of $S_3$}{$x=v$}
        \uIf{$p$ is first in $L$}{
            $F_g[x] = p$,\ $F_p[p] = s$\;
        }
        \Else($p$ is the last in $L$){
            $F_g[s] = p,\ F_p[p] = x$\;
            \lIf{$F_g[x] = -1$}{$F_g[x] = p$} \nllabel{line:back_link43}
        }
    }
  }  
%  \If{$I[-1] = I[0]$}{
%    $x = F_p[I[-1]]$\;
%    \If{$F_g[x]]=I[-1]$}{$F_g[x]=I[0]$}
%    $F_g[s_2] = I[0]$\;
%    remove $I[-1]$
%  }
\end{algorithm}

Algorithm \ref{algo:colect_23_ip_triangle} computes all ICs on the boundary of $S_2$.  It passes through every side $s$ of the triangle $S_2$ 
and computes the line-tetrahedron intersection $L$.
If the side is just touching $S_3$ and we get $L$ with single IC. These ICs will be rediscovered again in Algorithm \ref{algo:collect_23_ip_edges} 
with better topological information, however this is not the case if the touched edge $e$ of $S_3$ is coplanar with $S_2$ and the IC is inside of $e$.
To this end we save the IC into separate list $J$ and skip filling of the connection tables.
In the regular case, we process each of the two ICs in $L$ (loop on line \ref{line:one side loop}). The IC $p$ is added to the list $I$,
vertices of $S_2$ added twice are merged at final stage. Then (line \ref{line:m4_object}) we identify the \nface $x\in M_3$ 
the point $p$ lies on and set the tables $F_g$, $F_p$. For the IC at the vertex of $S_2$ we set $x$ to that vertex regardless of its position on $S_3$.
The backward temporary link on the line \ref{line:back_link43} is used later in Algorithm \ref{algo:collect_23_ip_edges} to allow proper continuation 
of the IC if it lies on edge or at vertex of $S_3$. The condition at the same line deals with the case that two sides of $S_3$ intersect 
the same object $x\in M_3$, in such case the back-link is unnecessary and would overwrite the correct link set by previous IC.

@@

  \lFor{edge $e$ of $S_3$}{$L[e]$ = intersection($e, S_2$)}  \nllabel{line:12edges}
  \For{edge $e$ of $S_3$ with non-empty $L[e]$}{
    $p=L[e]$\;
    \uIf{ $p$ is inside $e$}{ 
        \nllabel{line:edge_faces}        
        $(f_0,\ f_1)$ = \edgefaces($e$)
    } 
    \Else($p$ at the vertex $v$ of $S_3$){       
            $(f_0,\ f_1)$ = \vertexfaces($v$,$L$)   \tcp*{Algorithm \ref{algo:vertex_faces}}
            \nllabel{line:mark_edges}
    }    
    \uIf{ $p$ is on boundary of $S_2$ }{
         $p$ lies on edge or at vertex $x \in M_3$\;
         $q=F_g[x]$\tcp*{$q$ is already computed $p$}
         \lIf{$F_p[q]=x$}{$F_p[q]=f_1$}
         \lElse{$F_g[f_0]=q$}
         $F_g[x]=-1$\tcp*{remove the backlink}
    }
    \Else{
         append $p$ to $I$\;
         $F_g[f_0]=p,\ F_p[p]=f_1$\tcp*{overwrite the backlink}
    }   
  }  
\end{algorithm}

Algorithm \ref{algo:collect_23_ip_edges} uses the line-triangle intersection algorithm for the edges of $S_3$ (line \ref{line:12edges}).
First, the intersection $L[e]$ is evaluated for every edge $e$. The loop produces ICs in the interior of $S_2$ and possibly those ICs with special position on vertex or edge of $S_3$ already computed 
in Algorithm \ref{algo:colect_23_ip_triangle}. Then we pass through once again skipping edges with none or degenerate $L[e]$.
For every IC $p=L[e]$ we first get its (generalized) faces that are before and after the IC on IP. For an IC inside the edge $e$
the function \emph{edge faces}, returns its adjacent faces $f_0$, $f_1$ (see Fig. \ref{fig:edge_faces}). 
Their order is given by the sign permuted inner products in 1d-2d intersection. The order of faces match 
the order of sides of IP if the sign is negative. If the sign is positive the function \emph{edge faces} returns face pair $(f_1, f_0)$.
\begin{figure}[!htb]
  \begin{center}        
    \includegraphics[width=0.3\textwidth]{\figpath edge_faces.pdf}
  \end{center}
  \caption{Order of faces adjacent to the oriented edge $e$ pointing up.}
  \label{fig:edge_faces}
\end{figure}

The function \emph{vertex faces} (Algorithm \ref{algo:vertex_faces}, described later)
returns a pair of generalized faces (face or edge) possibly adjacent to the IC $L[e]$ at the vertex $v$ of $S_3$.
On line \ref{line:mark_edges} we mark all edges coincident with the vertex, since if they have any other intersection corner they are degenerate and 
not processed anyway.

@@

@@

\begin{algorithm}
  \caption{2d-3d intersection, order the points}  
  \label{algo:order_23}

  \SetKw{To}{to}
  \DontPrintSemicolon
  \KwIn{all points in $I$}
  \KwOut{polygon in $K$ in correct order}

  \lIf{$\abs{I} < 3$}{ return $J$}
  \lElse{ return $I$ sorted according to connectivity in $F_g$ and $F_p$}
\end{algorithm}
\noteJB{Write correct algo with deduplication of vertices.}
